---
title: "Pizza Order"
output: html_notebook
---


Import Data
```{r}
library (readr)
df <- read.csv("pizza_sales.csv")
df
summary(df)
```
Total Revenue
```{r}
library(tidyverse)
Total_rev <- df %>% summarise(Total=sum(total_price)) %>% format(nsmall=2)
Total_rev

#
x<-sum(df$total_price)
x
```
Average order value (not correct due to duplicate orders)
```{r}
no_orders <-  nrow(df)
average_order_value <- as.numeric(Total_rev)/no_orders
average_order_value

#or
average_order_value2<- mean(df$total_price)
average_order_value2


```
Average order value for distinct values
```{r}
#df$order_id %>% unique() --Gives unique values 
no_orders2<- n_distinct(df$order_id)
average_order_value3<- as.numeric(Total_rev)/no_orders2
average_order_value3

#
avg_test <- x/no_orders2
avg_test
```
No of pizzas
```{r}
sum(df$quantity)
```
Total orders
```{r}
no_orders2<- n_distinct(df$order_id)
no_orders2
```
Average Pizza
```{r}
avg_pizza_per_order <- sum(df$quantity)/no_orders2
avg_pizza_per_order
```
Get day,month and time respectively
```{r}
df$day <- str_sub(df$order_date,1,2)
df$month <- str_sub(df$order_date,4,5)
df$time <- str_sub(df$order_time,1,2)
df$weekday <- weekdays(dmy(df$order_date))
```

Graphs1: Daily Trend for Total orders
```{r}
library(DT)
library(ggplot2)
daily_orders <- df %>%
group_by(weekday) %>%
    distinct(order_id) %>%
  summarise(count=n())
datatable(daily_orders)

#Friday has the highest sales 
ggplot(daily_orders,aes(weekday,count,fill=weekday))+
  geom_bar(stat="identity") 

#Add labels of data to ggplot
ggplot(daily_orders,aes(weekday,count,fill=weekday))+
  geom_bar(stat="identity") +
  geom_text(aes(label=count),vjust = -0.5)

#Sort days of week
df$weekday <- factor(df$weekday,levels=c( "Sunday","Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))
ggplot(daily_orders,aes(weekday,count,fill=weekday))+
  geom_bar(stat="identity") +
  geom_text(aes(label=count),vjust = -0.5)
```

Graphs2: Hourly trend for total orders
```{r}
hourly_orders <- df %>%
                group_by(time) %>%
                distinct(order_id) %>%
                summarise(Total=n())
datatable(hourly_orders)

ggplot(hourly_orders,aes(time,Total,fill=Total))+
        geom_bar(stat = "identity")
#line chart
ggplot(hourly_orders,aes(time,Total,group=1))+
   geom_line(color="blue") 
```
Percentage of sales by category
```{r}
sales_category <- df %>%
 group_by(pizza_category) %>%
  summarise(Total=sum(total_price))  %>%
  #mutate(ratio=paste0(round((Total/sum(Total) *100),2),"%"))  %>%
  mutate(ratio=round((Total/sum(Total) *100),2))  %>%
  #Removed the Total Column
  select(-Total) 
datatable(sales_category)
```
Pie Chart
```{r}
ggplot(sales_category, aes(x = "", y = ratio, fill = factor(pizza_category)))+ #Format colors
 geom_bar(stat="identity",width=1)+
  coord_polar(theta = "y",start=0)+
 scale_fill_manual(values = c("Blue", "Red", "Green", "Orange")) +
geom_text(aes(label = ratio),
            position = position_stack(vjust = 0.5))+
   guides(fill=guide_legend(title="Pizza category"))

  
```
```{r}


```

percentage of pizza by size
```{r}
pizza_sizes <- df %>%
  group_by(pizza_size) %>%
  summarise(Total=sum(total_price)) %>%
  mutate(pct = round(Total/sum(Total)*100,2))
datatable(pizza_sizes)
```
Pizza size piechart

```{r}
ggplot(pizza_sizes,aes(x="",y=pct,fill=factor(pizza_size)))+
  geom_bar(stat="identity",width=1)+
  coord_polar(theta = "y",start=0)+
scale_fill_manual(values = c("Blue", "Red", "Green", "Orange","yellow")) +
geom_text(aes(label = pct),
            position = position_stack(vjust = 0.5))+
   guides(fill=guide_legend(title="Pizza Size"))
  
```


Total pizza sold by category
```{r}
Total_pizzas<- df %>% group_by(pizza_category) %>%
  summarise(Total=sum(quantity))
datatable(Total_pizzas)
```
```{r}
library(highcharter) 
Total_pizzas %>%
  hchart(
    "funnel", hcaes(x = pizza_category, y = Total),
    name = "Fruit consumption"
  )

```


Top five best sellers by total pizza sold 
```{r}
total_piz <- df %>%
  group_by(pizza_name) %>%
  summarise(Total_Pizzas_Sold=sum(quantity)) %>%
  arrange(desc(Total_Pizzas_Sold)) %>%
  head(5) 
 
datatable(total_piz)
```

```{r}
ggplot(total_piz,aes(Total_Pizzas_Sold,pizza_name,fill=Total_Pizzas_Sold)) +
  geom_bar(stat="identity")+
 geom_text(aes(label=Total_Pizzas_Sold), vjust=0.25)
  
```

Bottom five worst sellers by total pizza sold 
```{r}
total_piz_wos <- df %>%
  group_by(pizza_name) %>%
  summarise(Total_Pizzas_Sold=sum(quantity)) %>%
  arrange(Total_Pizzas_Sold) %>%
  head(5)

datatable(total_piz_wos)
```
Graph
```{r}
ggplot(total_piz_wos,aes(Total_Pizzas_Sold,pizza_name,fill=Total_Pizzas_Sold))+
  geom_bar(stat="identity")
```

